---
title: "Import/clean"
output:
  html_document:
    df_print: paged
---

## Setup

```{r setup}
library(tidyverse)
library(janitor)
library(RSocrata)
library(lubridate)
```

## Setting up download URL

Getting data directly from the [data.texas.gov](https://data.texas.gov/dataset/Mixed-Beverage-Gross-Receipts/naix-2893) portal using the Socrata API. Drawing from at least five years of data from the Mixed Beverage Gross Receipts for establishments with an Austin local address. Data provided by Texas Comptroller of Public Accounts.

```{r url}
mixbev_base_url = 'https://data.texas.gov/resource/fp9t-htqh.json?'
start_date = '2016-01-31'
end_date = '2021-08-31'
city = 'AUSTIN'

download_url <-  paste(
  mixbev_base_url,
  # "$limit=100&", 
  "$where=obligation_end_date_yyyymmdd%20between%20",
  "'", start_date, "'",
  " and ",
  "'", end_date, "'",
  "&location_city=",
  "'", city, "'",
  sep = ""
)

download_url
```

## Download the data

Downloading the data into an R object.

```{r download}
receipts_api <- read.socrata(download_url)

# look at the data
receipts_api %>% glimpse()
```

## Saving raw data into data-raw folder

Saving as .rds file

```{r raw}
receipts_api %>% write_rds("data-raw/01-mixbev.rds")
```


## Clean data and fix some values 

Changing the receipts columns from characters to numbers

```{r clean}
receipts_converted <- receipts_api %>%
  type_convert(
    cols(
        .default = col_character(), # sets a default of character unless specified below
        liquor_receipts = col_double(),
        wine_receipts = col_double(),
        beer_receipts = col_double(),
        cover_charge_receipts = col_double(),
        total_receipts = col_double()
    )
  )

receipts_converted %>% glimpse()
```

## Get helpful variables from dates

Changed obligation end date to date format. Parse the obligation end date into year, month, and year-month variables. Reordered and dropped some columns.

```{r parse-date}
date_fix <- receipts_converted %>% 
  mutate(
      obligation_end_date_yyyymmdd = ymd(obligation_end_date_yyyymmdd),
      yr = year(obligation_end_date_yyyymmdd),
      mo = month(obligation_end_date_yyyymmdd, label = TRUE),
      yr_mo = format(obligation_end_date_yyyymmdd, "%Y-%m")
  ) %>% 
  select(location_name, location_address, location_zip, obligation_end_date_yyyymmdd, yr, mo, yr_mo, total_receipts, everything(), -cover_charge_receipts, -location_state, -tabc_permit_number, -taxpayer_county, -location_number, -location_city)


date_fix %>% glimpse()
```

## Export the data

For later use in analysis notebook

```{r export}
date_fix %>% write_rds("data-processed/01-mix-bev.rds")
```





