---
title: "Mixed Beverage Analysis"
output:
  html_document:
    df_print: paged
---

## Goals

- Plot seasonal alcohol sales trends for downtown Austin
- Compare what a normal March year looked like versus during pandemic
- Calculate what percentage March sales make up of establishments’ yearly sales
- Find what downtown establishment had the most March alcohol sales over the 5 years

## Setup

```{r setup}
library(tidyverse)
library(lubridate)
library(scales)
library(plotly)
library(clipr)
library(dplyr)
```


## Import

Importing from previous notebook.

```{r import}
mix_bev <- read_rds("data-processed/01-mix-bev.rds")

mix_bev %>% glimpse()
```

## Filtering to downtown data

Only locations with 78701 zip code.

```{r filter-downtown}
downtown_data <- mix_bev %>% 
  filter(location_zip == "78701")

downtown_data
```
## Explore year-monthly total alcohol sales for downtown establishments

Was curious to see which year and month had the most alcohol sales.
_Answer: March 2019_

```{r yr-mo}
downtown_yrmo <- downtown_data %>% 
  group_by(yr_mo) %>% 
  summarize(
    total_receipts = sum(total_receipts)
    ) %>% 
  arrange(total_receipts %>% desc())

downtown_yrmo
```

## Find yearly total alcohol sales for downtown establishments

Wanted a summary table with total alcohol sales per year. Will need for calculating percentages in Question 3.

```{r yr-sales}
downtown_yr <- downtown_data %>% 
  group_by(yr) %>% 
  summarize(
    yr_total_receipts = sum(total_receipts)
    ) %>% 
  arrange(yr_total_receipts %>% desc()) %>% 
  mutate(yr = as.character(yr))

downtown_yr
```

## Question 1: Plot seasonal alcohol sales trends for downtown Austin

First graph is just to _explore_ the data and see seasonal trends.

```{r exploratory-plot, fig.width = 7}
downtown_yrmo %>% 
ggplot(aes(x = yr_mo, y = total_receipts)) +
  geom_line(group = 1) +
    theme(axis.text.x = element_text(angle = 90)) +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Seasonal Alcohol Sales Trends in Downtown Austin",
    x = "Month & Year",
    y = "Total Alcohol Sales" 
  )

```

## Split graph by year instead

Since the first graph is not easy to read, I split by year instead. In order to do this, I first prepared the data with month and year.

```{r seasonal-data}
yr_mo_data <- downtown_data %>% 
  group_by(yr, mo) %>% 
  summarize(
    total_receipts = sum(total_receipts)
  )

yr_mo_data
```

## Plot Seasonal Trends Split by Year

This is one of my two publishable graphs. It is now easier to tell that March leads downtown alcohol sales.

```{r seasonal-plot}
yr_mo_data %>% 
  ggplot(aes(x = mo, y = total_receipts)) +
  geom_line(aes(group = yr, color = as.character(yr))) +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Seasonal Alcohol Sales Trends in Downtown Austin, Split by Year",
    subtitle = "Monthly data from establishments selling liquor-by-the-drink (January 2016 to August 2021)",
    caption = "Mixed Beverage Gross Receipts data provided by Texas Comptroller of Public Accounts",
    x = "Month",
    y = "Total Alcohol Sales",
    color = "Year"
  )
  
```

## Question 2: Compare what a normal March year looked like versus during pandemic

Needed a summary table for March alcohol sales over the years. Had to add a third column for year in order to answer Question 3.

```{r march}
downtown_march <- downtown_data %>% 
  filter(mo == "Mar") %>% 
   group_by(yr_mo) %>% 
  summarise(
    mar_total_receipts = sum(total_receipts)
    ) %>% 
  arrange(mar_total_receipts %>% desc()) %>% 
  mutate(yr = str_sub(yr_mo, 1, 4))

downtown_march
```

## Get data for Datawrapper

My second publishable chart will be created using Datawrapper. Will create a column chart comparing March alcohol sales from 2016 to 2021.

```{r datawrapper}
downtown_march %>% write_clip(allow_non_interactive = TRUE)
```

[Here is my chart from Datawrapper.](https://www.datawrapper.de/_/MI67t/)

## Data preparation for Question 3

Joining year and year/month tables in order to calculate March percentages.

```{r join}
downtown_join <- downtown_yr %>% 
  left_join(downtown_march, by = "yr")

downtown_join
```

## Question 3: Calculate what percentage March sales make up of establishments’ yearly sales

Using joined table to calculate how much March contributes to sales of alcohol downtown during SXSW. I decided not to include the actual percentages in the data drop because they were lower than I expected, and it didn't sound as newsworthy or striking as I thought.

```{r percentages}
march_rates <- downtown_join %>% 
  mutate(
    march_percentage = ((mar_total_receipts / yr_total_receipts) * 100) %>% round()
  )

march_rates
```

## Question 4: Find  what downtown establishment had the most March alcohol sales over the 5 years

Answer: JW Marriott located by Austin Convention Center. Grouping by location name and location address to find exact establishment.

```{r most-march-sales}
downtown_data %>% 
  filter(mo == "Mar") %>% 
  group_by(location_name, location_address) %>% 
  summarize(
    march_total_receipts = sum(total_receipts)
    ) %>% 
  arrange(march_total_receipts %>% desc())
```

## Bonus: The Dogwood sales

From March 2019 to March 2020, 82% drop in sales. No sales generated in April or July of 2020.

```{r dogwood}
downtown_data %>% 
  filter(
    location_name == "THE DOGWOOD",
    location_address == "715 W 6TH ST"
    ) %>% 
  group_by(yr_mo) %>% 
  summarise(
    mo_total_receipts = sum(total_receipts)
    ) %>% 
  arrange(mo_total_receipts %>% desc())
```


## Dogwood March sales

Just looking at The Dogwood's alcohol sales during March

```{r dogwood-march}
downtown_data %>% 
  filter(
    mo == "Mar",
    location_name == "THE DOGWOOD",
    location_address == "715 W 6TH ST"
    ) %>% 
  group_by(yr_mo) %>% 
  summarise(
    mo_total_receipts = sum(total_receipts)
    ) %>% 
  arrange(mo_total_receipts %>% desc())
```

